Machine learning topological phases in real space

### 2020-07-24 07_paper/00_article/article.tex
Possible version to be resubmitted to prb

The main motivations for developing a data-driven approach based on real space are that wavevector space computations of topological invariants are only possible for systems with translational symmetry, which many physical systems of current interest (e.g. disordered systems in condensed matter) do not have. Moreover, since real space and wavevector space eigenvectors are related by Fourier transforms, the latter are essentially delocalized and therefore so is any information recovered from them. The data-driven approach described in this article addresses these issues. In particular, we employ it to scrutinize the issue of localizability of information in topological condensed matter systems.

@@

Perhaps the fundamental consequence of interpreting the information entropy density function $\rho_S(x)$ as the probability distribution arising from an information entropy wave function $\psi_S(x)$ defined on a 1D manifold is that it reconciles the apparently conflicting notions of a local topological signal arising from a global property of the SSH system. Indeed, while $\psi_S(x)$ is defined locally at every point of the 1D manifold, it is a single, global wave function encoding the distribution of topological information in a many-particle system. Therefore, the information entropy wave function $\psi_S(x)$ can be seen as an emergent property of a quantum many-body system.

@@

\bibliography{bibl}{}

>>>>

The main motivation for developing a data-driven approach based on real space is that the canonical method of choice for the analysis of topological systems, i.e. wavevector space computations of topological invariants, is often only feasible for systems with translational symmetry, which many physical systems of current interest (e.g. disordered systems in condensed matter) do not have. Furthermore, it is not always granted that the topological invariants of a physical system being investigated are known in advance, as is presently the case for many gapless insulators. In such cases, being able to engineer topological features that encode the topological states of a system while at the same time reducing the system's complexity may provide an alternative strategy. Moreover, since real space and wavevector space eigenvectors are related by Fourier transforms, the latter are essentially delocalized and therefore so is any information recovered from them. Constructing theoretical methods to trace the distribution of information in topological systems may be an essential prerequisite to discovering new topological invariants and features. The data-driven approach designed in this article addresses these issues. %In particular, we employ it to scrutinize the issue of localizability of information in topological condensed matter systems.

@@

Perhaps the fundamental consequence of interpreting the information entropy density function $\rho_S(x)$ as the probability distribution resulting from an information entropy wave function $\psi_S(x)$ defined on a 1D manifold is that it reconciles the apparently conflicting notions of a local topological signal arising from a global property of the SSH system. Indeed, while $\psi_S(x)$ is defined locally at every point of the 1D manifold, it is a single, global wave function encoding the distribution of topological information in a many-particle system. Therefore, the information entropy wave function $\psi_S(x)$ can be seen as an emergent property of a quantum many-body system.

@@

\bibliography{article_bib}{}

### 2020-07-24 07_paper/01_supplementary/supplementary.tex
Possible version to be resubmitted to prb

\begin{figure}
  \centering
  \includegraphics[width=.37\textwidth]{./phase_diagrams/ssh_model.pdf}
  \caption{Schematic set-up of a 1D SSH system. Here $t_1$ and $t_2$ are nearest-neighbor hoppings while $T_1$ and $T_2$ are second nearest-neighbor hoppings.}
\label{fig:model}
\end{figure}

\begin{figure}%[h!]
  \centering
  \includegraphics[width=.37\textwidth]{./phase_diagrams/winding.png}
  \caption{Winding number. The winding number of a closed, oriented curve with respect to a reference point is a topological invariant that counts how many times the curve winds around the point. Picture credits: Jim Belk, public domain.}
\label{fig:winding}
\end{figure}

@@

\begin{figure}
\centering
\subfigure[]{\label{eng_feat_ssh2:a}\includegraphics[width=.32\textwidth]{./supp_figures/ssh2/periodic_100_6561/real_space_best_12_sites/merge_imshow_winding_grids_sim.png}}
\subfigure[]{\label{eng_feat_ssh2:b}\includegraphics[width=.32\textwidth]{./supp_figures/ssh2/periodic_100_6561/dct_best_12_wavevector_sites_from_best_12_real_space_sites/merge_imshow_winding_grids_sim.png}}
\subfigure[]{\label{eng_feat_ssh2:c}\includegraphics[width=.32\textwidth]{./supp_figures/ssh2/periodic_100_6561/dst_best_12_wavevector_sites_from_best_12_real_space_sites/merge_imshow_winding_grids_sim.png}}
\caption{Phase diagrams learned from compressed representations of the SSH 2 system. (a) Phase diagram learned using the real space features \xSTwo.  (b) Phase diagram learned from the DCT topological features \xcSETwo. (b) Phase diagram learned from the DST topological features \xsSOTwo.}
\end{figure}

@@

It is clear from figures \ref{feature_importances_ssh1_longer_lattices} and \ref{feature_importances_ssh2_longer_lattices} that the patterns seen with 50 unit cells (figures 6(a) and 6(b) in the article) are stable across higher lattice sizes, with finer details emerging in the signals as the lattice size increases. This naturally leads us to speculation about the continuum limit of information entropy signatures, a matter formally explored in the article.

@@

\begin{thebibliography}{10}
\bibitem{asboth2016short}
J.~K. Asb{\'o}th, L.~Oroszl{\'a}ny, and A.~P{\'a}lyi, ``A short course on
  topological insulators,'' {\em Lecture notes in physics}, vol.~919, 2016.

\bibitem{goodfellow2016deep}
I.~Goodfellow, Y.~Bengio, and A.~Courville, {\em Deep learning}.
\newblock MIT press, Cambridge, MA, USA, 2016.

\bibitem{bishop2006pattern}
C.~M. Bishop, {\em Pattern recognition and machine learning}.
\newblock Springer New York, NY, USA, 2006.

\bibitem{cayton2005algorithms}
L.~Cayton, ``Algorithms for manifold learning,'' {\em Univ. of California at
  San Diego Tech. Rep}, vol.~12, no.~1-17, p.~1, 2005.

\bibitem{narayanan2010sample}
H.~Narayanan and S.~Mitter, ``Sample complexity of testing the manifold
  hypothesis,'' in {\em Advances in Neural Information Processing Systems},
  pp.~1786--1794, 2010.

\bibitem{rifai2011manifold}
S.~Rifai, Y.~N. Dauphin, P.~Vincent, Y.~Bengio, and X.~Muller, ``The manifold
  tangent classifier,'' in {\em Advances in Neural Information Processing
  Systems}, pp.~2294--2302, 2011.
\end{thebibliography}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=.37\textwidth]{./ssh/ssh_model.pdf}
  \caption{Schematic set-up of a 1D SSH system. Here $t_1$ and $t_2$ are nearest-neighbor hoppings while $T_1$ and $T_2$ are second nearest-neighbor hoppings.}
\label{fig:model}
\end{figure}

\begin{figure}%[h!]
  \centering
  \includegraphics[width=.37\textwidth]{./ssh/winding.png}
  \caption{Winding number. The winding number of a closed, oriented curve with respect to a reference point is a topological invariant that counts how many times the curve winds around the point. Picture credits: Jim Belk, public domain.}
\label{fig:winding}
\end{figure}

@@

\begin{figure}
\centering
\subfigure[]{\label{eng_feat_ssh2:a}\includegraphics[width=.32\textwidth]{./supp_figures/ssh2/periodic_100_6561/real_space_best_12_sites/merge_imshow_winding_grids_sim.png}}
\subfigure[]{\label{eng_feat_ssh2:b}\includegraphics[width=.32\textwidth]{./supp_figures/ssh2/periodic_100_6561/dct_best_12_wavevector_sites_from_best_12_real_space_sites/merge_imshow_winding_grids_sim.png}}
\subfigure[]{\label{eng_feat_ssh2:c}\includegraphics[width=.32\textwidth]{./supp_figures/ssh2/periodic_100_6561/dst_best_12_wavevector_sites_from_best_12_real_space_sites/merge_imshow_winding_grids_sim.png}}
\caption{Phase diagrams learned from compressed representations of the SSH 2 system. (a) Phase diagram learned using the real space features \xSTwo.  (b) Phase diagram learned from the DCT topological features \xcSETwo. (b) Phase diagram learned from the DST topological features \xsSOTwo.}
\label{eng_feat_ssh2}
\end{figure}

@@

It is clear from figures \ref{feature_importances_ssh1_longer_lattices} and \ref{feature_importances_ssh2_longer_lattices} that the patterns seen with 50 unit cells (figures 6(a) and 6(b) in the article) are stable across higher lattice sizes, with finer details emerging in the signals as the lattice size is increased. This naturally leads us to speculation about the continuum limit of information entropy signatures, a matter formally explored in the article.

@@

\bibliography{supplementary_bib}{}

### 2020-07-24 10_prb_resubmission_01/00_article/article.tex
Possible version to be resubmitted to prb

%\documentclass[aps,prb,amsmath,twocolumn,amssymb,titlepage]{revtex4-1}
\documentclass[10pt]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{nicefrac}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath} 
\usepackage{subfigure}
\usepackage{multirow} 
\usepackage{tabularx} 
\usepackage{array}
\usepackage{units}
\usepackage{tensor} 
\usepackage{braket}
\usepackage{bm}
\usepackage{hyperref}
\usepackage[resetlabels, labeled]{multibib}
\usepackage{xcolor}
%\newcites{supp}{References}

@@

The main motivations for developing a data-driven approach based on real space are that wavevector space computations such as eq. \eqref{winding_number} are only possible for systems with translational symmetry, which many physical systems of current interest (e.g. disordered systems in condensed matter) do not have. Moreover, since real space and wavevector space eigenvectors are related by Fourier transforms, the latter are essentially delocalized and therefore so is any information recovered from them. We argue here that a data-driven approach based on real space offers a viable alternative to wavevector space computations that addresses these shortcomings.

>>>>

%\documentclass[aps,prb,amsmath,twocolumn,amssymb,titlepage]{revtex4-1}
\documentclass[10pt]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{nicefrac}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath} 
\usepackage{subfigure}
\usepackage{multirow} 
\usepackage{tabularx} 
\usepackage{array}
\usepackage{units}
\usepackage{tensor} 
\usepackage{braket}
\usepackage{bm}
\usepackage{hyperref}
\usepackage[resetlabels, labeled]{multibib}
%\newcites{supp}{References}

@@
